---
title: "Applying resampling methods to regression model"
output: html_notebook
---

```{r}
# importing packages
library(dplyr)
library(fastDummies)
library(ggplot2)
library(tidyr)
library(diptest)
library(LaplacesDemon)
library(mousetrap)
library(multimode)
library(haven)
library(datasets)
library(jtools)
library(car)
library(caret)
library(psych)
library(rlang)
library(boot)
```
```{r}
# reading data
df <- read.csv("https://raw.githubusercontent.com/meetgojiya98/Business_Analytics_MBA6636/main/Assignment%201/MBA6636_SM21_Professor_Proposes_Data.csv")
```
```{r}
#changing the type of 'Price' variable to numeric type and removing the '$' sign
df$Price = as.numeric(gsub('[$,]', '', df$Price))
```
```{r}
#Filling in missing values 
df <- df %>% fill(Price, .direction = 'down')
```
```{r}
# Clubbing the 'I2' clarity type into 'I1', 'VVS1' and 'VVS2' clarity type into 'VS2' and creating a new variable
df <- df %>%
  mutate(ClarityN = case_when(
    startsWith(Clarity, "I1") ~ "I1",
    startsWith(Clarity, "I2") ~ "I1",
    startsWith(Clarity, "VVS1") ~ "VS2",
    startsWith(Clarity, "VVS2") ~ "VS2",
    startsWith(Clarity, "SI1") ~ "SI1",
    startsWith(Clarity, "SI2") ~ "SI2",
    startsWith(Clarity, "SI3") ~ "SI2",
    startsWith(Clarity, "VS1") ~ "VS1",
    startsWith(Clarity, "VS2") ~ "VS2"
    ))
```
```{r}
# Clubbing the colours into 4 categories and creating a new variable
df <- df %>%
  mutate(ColourN = case_when(
    startsWith(Colour, "D") ~ "Colorless",
    startsWith(Colour, "E") ~ "Colorless",
    startsWith(Colour, "F") ~ "Colorless",
    startsWith(Colour, "G") ~ "Near Colorless",
    startsWith(Colour, "H") ~ "Near Colorless",
    startsWith(Colour, "I") ~ "Near Colorless",
    startsWith(Colour, "J") ~ "Faint Yellow",
    startsWith(Colour, "K") ~ "Faint Yellow",
    startsWith(Colour, "L") ~ "Very Light Yellow"
    ))
```
```{r}
# Clubbing the colours into 2 categories and creating a new variable
df <- df %>%
  mutate(Colour2C = case_when(
    startsWith(ColourN, "Colorless") ~ "Colorless",
    startsWith(ColourN, "Near Colorless") ~ "Colorless",
    startsWith(ColourN, "Faint Yellow") ~ "Yellow",
    startsWith(ColourN, "Very Light Yellow") ~ "Yellow"
    ))
```
```{r}
# Creating new variable based off of the Cut variable
df <- df %>%
  mutate(CutN = case_when(
    startsWith(Cut, "F") ~ "Fair",
    startsWith(Cut, "G") ~ "Good",
    startsWith(Cut, "I") ~ "Ideal",
    startsWith(Cut, "V") ~ "Very Good",
    startsWith(Cut, "X") ~ "Excllent"
    ))
```
```{r}
# Clubbing the Certifications into 2 categories and creating a new variable
df <- df %>%
  mutate(CertG = case_when(
    startsWith(Certification, "AGS") ~ "AGS - GIA",
    startsWith(Certification, "DOW") ~ "Rest",
    startsWith(Certification, "EGL") ~ "Rest",
    startsWith(Certification, "GIA") ~ "AGS - GIA",
    startsWith(Certification, "IGI") ~ "Rest"
    ))
```
```{r}
# Creating new variable based off of the Polish variable
df <- df %>%
  mutate(PolishN = case_when(
    startsWith(Polish, "F") ~ "Fair",
    startsWith(Polish, "G") ~ "Good",
    startsWith(Polish, "I") ~ "Ideal",
    startsWith(Polish, "V") ~ "Very Good",
    startsWith(Polish, "X") ~ "Excllent"
    ))
```
```{r}
# Creating new variable with 2 categories based off of the Polish variable
df <- df %>%
  mutate(Polish2C = case_when(
    startsWith(PolishN, "Fair") ~ "Fair - Good",
    startsWith(PolishN, "Good") ~ "Fair - Good",
    startsWith(PolishN, "Ideal") ~ "Very Good - Ideal",
    startsWith(PolishN, "Very Good") ~ "Very Good - Ideal",
    startsWith(PolishN, "Good") ~ "Very Good - Ideal",
    startsWith(PolishN, "Excllent") ~ "Very Good - Ideal"
    ))
```
**Polish**: The frequency of Ideal and Fair category within the Polish variable were extremely low. In order to bring balance, Polish variable is categorized in two categories, Fair - Good, and Very Good - Ideal.
```{r}
# Creating new variable based off of the Symmetry variable
df <- df %>%
  mutate(SymmetryN = case_when(
    startsWith(Symmetry, "F") ~ "Fair",
    startsWith(Symmetry, "G") ~ "Good",
    startsWith(Symmetry, "I") ~ "Ideal",
    startsWith(Symmetry, "V") ~ "Very Good",
    startsWith(Symmetry, "X") ~ "Excellent",
    ))
```
```{r}
# Creating high carat, high price dataset 
highPricedf <- df %>%
  filter(Price >= 1500) 
```
```{r}
# Creating low carat, low price dataset 
lowPricedf <- df %>%
  filter(Price < 1500) 
```
# Model Building
```{r}
# running a simple linear regression model with Price as a response variable and Carat as explanatory variable 
priceCarat <- lm(formula = Price ~ Carat, data = highPricedf)

# model summary
summary(priceCarat)
summ(priceCarat)
```
There is some correlation between Carat and Price variables as the R-squared is around 11%. But the regression coefficient of Carat is negative, which is unexpected as high carat diamonds must have high price. Carat has a p-value of 2.12e-07. Since this value is in the range [0, 0.001], it has a significance code of ***. Hence, Carat is statistically significant predictor of diamond price.
```{r}
# Plotting a simple linear regression model with Price as a response variable and Carat as explanatory variable 
effect_plot(priceCarat, pred = Carat, interval = TRUE, plot.points = TRUE, rug = TRUE, partial.residuals = TRUE)
```
```{r}
# running a simple linear regression model with Price as a response variable and ClarityN as explanatory variable 
priceClarityN <- lm(formula = Price ~ ClarityN, data = df)

# model summary
summary(priceClarityN)
summ(priceClarityN)
```
The R-squared is reasonably high at around 30%, which indicated strong correlation between Price and Clarity. Since clarity's p-value value is in the range [0, 0.001], it has a significance code of ***. Hence, Carat is statistically significant predictor of diamond price.
```{r}
# Plotting a simple linear regression model with Price as a response variable and ClarityN as explanatory variable 
effect_plot(priceClarityN, pred = ClarityN, interval = TRUE, plot.points = TRUE, rug = TRUE, partial.residuals = TRUE)
```
```{r}
# running a simple linear regression model with Price as a response variable and CertG as explanatory variable 
priceCertG <- lm(formula = Price ~ CertG, data = highPricedf)

# model summary
summary(priceCertG)
summ(priceCertG)
```
After clubbing the categories, we observed that the R-squared is around 5%, which is noticeable.
```{r}
# Plotting a simple linear regression model with Price as a response variable and CertG as explanatory variable 
effect_plot(priceCertG, pred = CertG, interval = TRUE, plot.points = TRUE, rug = TRUE, partial.residuals = TRUE)
```
```{r}
# running a simple linear regression model with Price as a response variable and CutN as explanatory variable 
priceCutN <- lm(formula = Price ~ CutN, data = highPricedf)

# model summary
summary(priceCutN)
summ(priceCutN)
```
Here, the R-suared is 14%, which is considered as good.It also has high significance level. So, overall it's a good feature.
```{r}
# Plotting a simple linear regression model with Price as a response variable and CutN as explanatory variable 
effect_plot(priceCutN, pred = CutN, interval = TRUE, plot.points = TRUE, rug = TRUE, partial.residuals = TRUE)
```
```{r}
# running a simple linear regression model with Price as a response variable and Polish2C as explanatory variable 
pricePolish2C <- lm(formula = Price ~ Polish2C, data = highPricedf)

# model summary
summary(pricePolish2C)
summ(pricePolish2C)
```
The R-squared is 11%, which indicates good correlation between Price and Polish. Polish is statistically significant predictor of diamond price.
```{r}
# Plotting a simple linear regression model with Price as a response variable and Polish2C as explanatory variable 
effect_plot(pricePolish2C, pred = Polish2C, interval = TRUE, plot.points = TRUE, rug = TRUE, partial.residuals = TRUE)
```
```{r}
# running a simple linear regression model with Price as a response variable and SymmetryN as explanatory variable 
priceSymmetryN <- lm(formula = Price ~ SymmetryN, data = df)

# model summary
summary(priceSymmetryN)
summ(priceSymmetryN)
```
The R-squared is just around 6%, which indicated not so good correlation between Price and Symmetry. Since Since it has no significance code, it's not a statistically significant predictor of diamond price.
```{r}
# Plotting a simple linear regression model with Price as a response variable and SymmetryN as explanatory variable 
effect_plot(priceSymmetryN, pred = SymmetryN, interval = TRUE, plot.points = TRUE, rug = TRUE, partial.residuals = TRUE)
```
```{r}
# running a simple linear regression model with Price as a response variable and Colour2C as explanatory variable 
priceColour2C <- lm(formula = Price ~ Colour2C, data = highPricedf)

# model summary
summary(priceColour2C)
summ(priceColour2C)
```
R-square is just around 2%, which is low. It a sign of weak correlation.
```{r}
# Plotting a simple linear regression model with Price as a response variable and Colour2C as explanatory variable 
effect_plot(priceColour2C, pred = Colour2C, interval = TRUE, plot.points = TRUE, rug = TRUE, partial.residuals = TRUE)
```
```{r}
# running a multiple linear regression model with Price as a response variable and CutN, Colour2C, ClarityN, CertG, Polish2C, Carat as explanatory variable 
priceMulti <- lm(formula = Price ~ CutN + Colour2C + ClarityN + CertG + Polish2C + Carat, data = highPricedf)

# model summary
summary(priceMulti)
summ(priceMulti)
```
Here, R-squared is 45% and adjusted R-squared is 42%, which is fairly high. This model is statistically significant predictor of the price of the diamond.

```{r}
# running a multiple linear regression model with Price as a response variable and Colour2C, ClarityN, CertG, Polish2C, Carat as explanatory variable 
priceMulti2 <- lm(formula = Price ~ Colour2C + ClarityN + CertG + Polish2C + Carat, data = lowPricedf)

# model summary
summary(priceMulti2)
summ(priceMulti2)
```
Here, the R-squared is extremely high, coming at 93%. 

# Cross-validation
```{r}
df_ctrl <- trainControl(method = "cv", number = 5)

# 5-Fold cross-validation
priceMultiCv <- lm(formula = Price ~ CutN + Colour2C + ClarityN + CertG + Polish2C + Carat, data = highPricedf)
priceMultiCv
# model summary
summary(priceMultiCv)
summ(priceMultiCv)
```
```{r}
df_ctrl <- trainControl(method = "cv", number = 3)

# 3-Fold cross-validation
priceMultiCv <- lm(formula = Price ~ CutN + Colour2C + ClarityN + CertG + Polish2C + Carat, data = highPricedf)
priceMultiCv
# model summary
summary(priceMultiCv)
summ(priceMultiCv)
```
```{r}
# Fit lm model using 10-fold CV: model
looMo <- train(
  Price ~ CutN + Colour2C + ClarityN + CertG + Polish2C + Carat, 
  highPricedf,
  method = "lm",
  trControl = trainControl(
    method = "cv", 
    number = 10,
    verboseIter = TRUE
  ),
  na.action = na.pass
)

# Print model to console
looMo
```

```{r}
#Leave-One-Out Cross-Validation
loo <- trainControl(method = "LOOCV")

#fit a regression model and use LOOCV to evaluate performance
looModel <- train(Price ~ CutN + Colour2C + ClarityN + CertG + Polish2C + Carat, data = highPricedf, method = "lm", trControl = loo, na.action = na.pass)

#view summary of LOOCV               
print(looModel)
```
# Bootstrap
```{r}
# Bootstrap resampling
bootFun <- function(d, i){
	d2 <- d[i,]
	return(cor(highPricedf$Price, highPricedf$Carat))
}

set.seed(626)
bootcr <- boot(highPricedf, bootFun, R=500)
bootcr
summary(bootcr)
```











































